---
title: "Quick Project 1 Technical Appendix"
author: "Jordan Gropper"
date: "7/7/21"   
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

# set up data, tidy data, and EDA

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE}
# Course: 5210 Communicating Data
# Purpose: Quick Project 1 Technical Appendix
# Date: July 7th, 2021
# Author: Jordan Gropper

```

```{r message = FALSE}
library(ggplot2) 
library(gridExtra) # use to put graphs together in the same frame
library(janitor)
library(knitr)
library(GGally)
library(scales) # to get $ and , on axis in graph
library(tidyverse)
library(corrplot)#used for correlation plot with good looks below
```

# Load and transform data

```{r}
houses <- read.csv("qp1_data.csv")#read data into environment and assign to variable

houses <- houses %>% 
  mutate(living_sqft_price = price/sqft_living)

houses <- houses %>% 
  mutate(lot_sqft_price = price/sqft_lot)
#class(living_sqft_price)
```


# Base EDA Step 1: Uni-variate non-graphical EDA
```{r}
head(houses, 10)#check first ten
str(houses)#check structure
```

+ EDA findings non graphical:

    - The data consists of 13 original well organized variables. These certainly make good sense in the context of looking into
    the housing market.
    -I believe that it would be useful to manipulate the data into another column or columns for meausuring single ratios. Such
      as prcie per square living space foot and price per square foot of lot
    - The data is Tidy with each observation(each home) being in one row, and each variable has its own column.with 21,613           observations we have a lot of good information to go off of.

```{r}
#Summary check for the data sets statistics
summary(houses)#check sum stats.
```
    
**Comments**
+ Summary stats on Quantitative

    - Price and Bedrooms are not symetric as they are both skewed to the right with the mean being 
      greater than the median by $90,000 for price, and by .37 of a room for bedrooms. 
        - This indicates that median will be a better representation than the mean
        -This works well for bedrooms too as it would be best to not observe fractions of a room
        - Additionally it means that we will have to verify assumptions about the ability to 
          linearly model this data if we are so inclined to produce a model.
        - There could be a few outliers in this data set for price and bedroom.
            - The max for both are well outside the 3rd quartile range of the data
    - For the ratios of price per sqft of lot and living space they also have gigantic outliers, where 
    the price per foot of space is multitudes larger than the 3rd quartile.
+ Summary stats of Categorical

    - The included variables of yr_renovated and waterfront seem to vastly say that the majority of the homes
      don't have a waterfront view and have not been renovated.
    - The same could be said for the "grade" of the home
        - In the context of the data both bedrooms and bathrooms are likely to have a few large outliers.
    - The zip codes are relatively irrelevant for this type of analysis.
    -It would be a good idea to take off the top and bottom 10% of some observations in final stages to observe
    trends without bias.

**Questions**

+ Data is unbalanced in many of the categorical variables

    - are these variables still useful, as they have almost all observations in on categorey
+ Skew is heavy in some of the quantitative variables
    
    - Should some of the extreme outliers be removed?
    -are there other ways we will want to be cautious with the skew,
      beyond focusing on median instead of mean?
        - ouliers could distort descriptive statistics
    - check all uni-variate non-graphical findings visually
    
# Base EDA Step 2: Uni-variate graphical EDA

+ Here we will examine each variable individually

## Categorical/Factor variables:

```{r}
watrfrnt <- ggplot(data = houses, mapping = aes(waterfront))
view_grade <- ggplot(data = houses, mapping = aes(view))
build_qaulity <- ggplot(data = houses, mapping = aes(grade))
current_cond <- ggplot(data = houses, mapping = aes(condition))
year_built <- ggplot(data = houses, mapping = aes(yr_built))
yr_reno <- ggplot(data = houses, mapping = aes(yr_renovated))
zip_code <- ggplot(data = houses, mapping = aes(zip_code))

watrfrnt + geom_bar(fill = "blue")
view_grade + geom_bar(fill = "blue")
build_qaulity + geom_bar(fill = "blue")
current_cond + geom_bar(fill = "blue")
year_built + geom_bar(fill = "blue")
yr_reno + geom_bar(fill = "blue", binwidth = 1000) 
zip_code + geom_bar(fill = "blue")
```
    
**Comments**

+ The majority of homes do not have a water front view

    - it appears we may have a lot of inner city rural homes in this data

+ Most homes are in moderate condition with a rating of 3

    - Many are in reasonable condition, and were built at a moderate grade

+ The majority of the homes have never been renovated

    - renovations are mostly nonexistant, just general maintenance performed.

## Quantitative variables

### Price   
```{r}
price <- ggplot(data = houses, mapping = aes(price))
price + geom_histogram(fill = "blue") + scale_x_continuous(labels = scales :: dollar)
```

### Bedrooms
```{r}
bedrooms <- ggplot(data = houses, mapping = aes(bedrooms))
bedrooms + geom_histogram(fill = "blue")
```

### Bathrooms
```{r}
bathrooms <- ggplot(data = houses, mapping = aes(bathrooms))
bathrooms + geom_histogram(fill = "blue")
```

### Square Feet of living space
```{r}
sqft_living <- ggplot(data = houses, mapping = aes(sqft_living))
sqft_living + geom_histogram(fill = "blue")
```

### Size of Lot
```{r}
sqft_lot <- ggplot(data = houses, mapping = aes(sqft_lot))
sqft_lot + geom_histogram(fill = "blue")
```

### Number of Stories
```{r}
levels <- ggplot(data = houses, mapping = aes(floors))
levels + geom_histogram(fill = "blue") 
```
```{r}
per_sqft_living <- ggplot(data = houses, mapping = aes(living_sqft_price))
per_sqft_living + geom_histogram(fill = "blue") +
  scale_x_continuous(labels = scales :: dollar)

```
```{r}
per_sqft_lot <- ggplot(data = houses, mapping = aes(lot_sqft_price))
per_sqft_lot + geom_histogram(fill = "blue") +
  scale_x_continuous(labels = scales :: dollar)

```
+ Portrays the variety of skew for each variable
+ Quantitative variables Price, Bedrooms, and sqft of living all have very harsh fall off, and long tail to right.

    - will use median rather than mean for analysis
    - This portrays how there are most likely some "mansions" in the data set causing skew.
    - but they are important to our research question so will not remove them at this point.
    - It would be a good idea to take the top 10% and bottom 10% of observations off to check at some point
+ Bathrooms is the variable that seems the most normally distributed

    - Even here though there is the oddity of bathrooms being counted as a fraction of a unit.
+ The Mutated variables I added on are also skewed with long tails to the right.

    - This makes sense as houses with extrodinarily high price per sqft of living and 
    of the lot are uncommon.
    
### Full Categorical
```{r  echo = FALSE}
grid.arrange(
  
watrfrnt + geom_bar(fill = "blue"),

view_grade + geom_bar(fill = "blue"),

build_qaulity + geom_bar(fill = "blue"),

current_cond + geom_bar(fill = "blue"),

year_built + geom_bar(fill = "blue"),

yr_reno + geom_bar(fill = "blue", binwidth = 10),

zip_code + geom_bar(fill = "blue"),

ncol = 2 )
```

### Full Quantitative
```{r  echo = FALSE}
grid.arrange(
#price <- ggplot(data = houses, mapping = aes(price))
price + geom_histogram(fill = "blue") + scale_x_continuous(labels = scales :: dollar), 

#bedrooms <- ggplot(data = houses, mapping = aes(bedrooms))
bedrooms + geom_histogram(fill = "blue"), 

#bathrooms <- ggplot(data = houses, mapping = aes(bathrooms))
bathrooms + geom_histogram(fill = "blue"), 

#sqft_living <- ggplot(data = houses, mapping = aes(sqft_living))
sqft_living + geom_histogram(fill = "blue"), 

#sqft_lot <- ggplot(data = houses, mapping = aes(sqft_lot))
sqft_lot + geom_histogram(fill = "blue"), 

#levels <- ggplot(data = houses, mapping = aes(floors))
levels + geom_histogram(fill = "blue"),

per_sqft_living + geom_histogram(fill = "blue"),

per_sqft_lot + geom_histogram(fill = "blue"),

ncol = 2 )
```

# Base EDA Step 3: Multi-variate non-graphical 

## Categorical 

### Grade and view

```{r}
# Countingency/cross table
houses %>% 
  tabyl(grade, view) %>% # creates table of counts
  adorn_totals(where = c("row", "col")) # Total margins

```

```{r}
# Proportion contingency/cross table
houses %>% 
  tabyl(grade, view) %>% # creates table of counts
  adorn_totals(where = c("row", "col")) %>%  # Total margins
  adorn_percentages(denominator = "all") %>% # creates proportions
  adorn_rounding(2) # round decimals

```
### Grade and Condition
```{r}
# Proportion contingency/cross table
houses %>% 
  tabyl(grade, condition) %>% # creates table of counts
  adorn_totals(where = c("row", "col"))

```


```{r}
# Proportion contingency/cross table
houses %>% 
  tabyl(grade, condition) %>% # creates table of counts
  adorn_totals(where = c("row", "col")) %>%  # Total margins
  adorn_percentages(denominator = "all") %>% # creates proportions
  adorn_rounding(2) # round decimals

```

### Zip code and Year Renovated
```{r}
# Proportion contingency/cross table
houses %>% 
  tabyl(zip_code, yr_renovated) %>% # creates table of counts
  adorn_totals(where = c("row", "col")) 
```
```{r}
# Proportion contingency/cross table
houses %>% 
  tabyl(zip_code, yr_renovated) %>% # creates table of counts
  adorn_totals(where = c("row", "col")) %>%  # Total margins
  adorn_percentages(denominator = "all") %>% # creates proportions
  adorn_rounding(2) # round decimals

```
### Year Built and Current Condition
```{r}
# Proportion contingency/cross table
houses %>% 
  tabyl(yr_built, condition) %>% # creates table of counts
  adorn_totals(where = c("row", "col")) 
```

```{r}
# Proportion contingency/cross table
houses %>% 
  tabyl(yr_built, condition) %>% # creates table of counts
  adorn_totals(where = c("row", "col")) %>%  # Total margins
  adorn_percentages(denominator = "all") %>% # creates proportions
  adorn_rounding(2) # round decimals

```
**Comments**

+ shows that the grades are lumped near the middle, with majority having a low view rating.

    - This means that the majority of homes are just built moderately well in our data
    - Indicates that a home doesn't have to be built perfectly to then sell well later

+ There definitely seems to be a trend between initial build quality, and current condition.

    - Those that are lumped within the medium range of build quality, tend to stay in the mid range.
    - It is unlikely that a home will increase its condition over the years from its initial build grade
    it appears so far, will look into more.
+ There is not a significant amount of renovated homes. When put into rounded proportions, none are shwon.

    - It may be a good idea to remove the renovation variable
+ Build quality is spread out over build time

    - It does not guarantee that just because a home is older that it will be in poor condition.

**Questions**

+ Need to check one overlap of condition, grade and price

+ should we remove renovate since it is so unuseful?


## Quantitative

```{r}
#uses all my variables? SHould I code it to be just the first 6?
houses %>% 
  select_if(is.numeric) %>% # Use to select just the numeric variables
  cor() %>% 
  round(2) %>% 
  kable()
```

**Comments**

+ Strong correlation between price and sqft_living, as well as the initial grade of the build
  
    - This makes good sense, but it is interesting that it appears that current condition has little 
    to do with price.
    - Also interesting that price doesn't seem to correlate with size of the lot or the year it was built
+ Not surprising That something like zipcode has little to no correlation on any of the variables.

+ There are no strong correlations for the price per foot calculations for lot and living space .
  
    - I would interpret this as there being no single variable that makes an amount of space worth something.
    -Each variable of the condition, grade, view and location all impact the price. 

**Questions**

+ Does positive correlation imply causation: The higher the price, the larger the house?

    - Makes sense since the larger the home, the more materials needed.


# Base EDA Step 4: Multi-variate graphical

## Step 4.1: Categorical

+ Bar graphs with multiple categorical variables


```{r}
condition <- factor("condition", levels = 1:5)
class(condition)#should have fixed the problem but it
condition
grid.arrange(
  houses %>% 
    ggplot(mapping = aes(x = yr_built, fill = as.factor(condition))) +
    geom_bar(position = "dodge"), 
  
  houses %>% 
    ggplot(mapping = aes(x = zip_code, fill = as.factor(condition))) +
    geom_bar(position = "dodge", mapping = aes(condition)), 
  
  houses %>% 
    ggplot(mapping = aes(x = view, fill = as.factor(condition))) +
    geom_bar(position = "dodge"),
  
  houses %>% 
    ggplot(mapping = aes(x = grade)) +
    geom_bar(position = "dodge", mapping = aes(fill = as.factor(condition))),
  
  ncol = 2)


```
```{r}
grid.arrange(
  # Cluster of counts
  houses %>% 
    ggplot(mapping = aes(x = yr_built)) + # Add additional variable by specify "fill"
      geom_bar(mapping= aes(fill = as.factor(condition))),
  # Proportion of counts
  houses %>% 
    group_by(yr_built, grade) %>% 
    ggplot(mapping = aes(x = yr_built, fill = as.factor(grade))) + # Add additional variable by specify "fill"
      geom_bar(mapping = aes(fill = as.factor(condition))) +
      coord_flip(),

ncol = 1
)
```
**Comments**
 
+ We have good spread of data on building years.

    - unusually low in the 2008 time period, maybe because of housing crash
    - We have more data on homes built more recently
    - The concentration of quality is spread out of the years.
    - older homes have more of a chance of decaying quality
    - more mid range quality homes built recently

**Questions**

+ How does price vary per year on the homes?

+ Do only old homes degrade?




## Step 4.2: Quantitative

```{r}
# Code scatterplots using grid.arrange so can see all quant variables of interest together
houses %>% 
  ggplot(mapping = aes(x = price, y = floors)) +
  geom_point() + scale_x_continuous(labels = scales :: dollar)

```
```{r}
# Code scatterplots using grid.arrange so can see all quant variables of interest together
houses %>% 
  ggplot(mapping = aes(x = price, y = bedrooms)) +
  geom_point() + scale_x_continuous(labels = scales :: dollar)

```
```{r}
# Code scatterplots using grid.arrange so can see all quant variables of interest together
houses %>% 
  ggplot(mapping = aes(x = price, y = sqft_living)) +
  geom_point() + scale_x_continuous(labels = scales :: dollar)

```
**Comments**

+ Interesting pattern

    - Bedrooms, space and floors usually all impact price
    - however there are multi floor houses with rooms that are cheaper then they should be
    - This could be due to quality and condition of the home
    
**Questions**

+ How do grade, condition and size interact to cause this disparity?

    - observe median price per square foot in comparison with build details

## Step 4.3: Categorical and quantitative

```{r message = FALSE}
# Use ggpairs from the GGally package
houses %>%
  ggpairs()

```

```{r}
cor_houses<- (cor(houses))
corrplot(cor_houses,
         method = "number",
         sig.level = 0.05,
         order = "original",
         diag = FALSE,
         type = "upper",
         tl.srt = 45,
         tl.col = "black")

```
**Comments**

+ Each factor contributes reasonably

    - both grade and size of the house impact the price
    -grade and condition are intertwine slightly
    - the size of the living space impacts initial grade
      - nice large homes

**Questions**

+ Look more closely at:
    
    - location, condition, grade, price, size
    - size and price together

## Location, price and quality

**Follow up on correlation graph questions below**


```{r}
# Show code
houses <- read.csv("qp1_data.csv")
#head(houses, 10)
str("condition")
grid.arrange(
  # age, price size and condition
  houses %>% 
    ggplot(mapping = aes(x = price, y = sqft_living, color = condition)) +
      geom_point() +
      geom_smooth(method = "lm", se = FALSE) +
    scale_x_continuous(labels = scales :: dollar),
  
    houses %>% 
    ggplot(mapping = aes(x = price, y = zip_code, color = grade)) +
      geom_point() +
      geom_smooth(method = "lm", se = FALSE) +
    scale_x_continuous(labels = scales :: dollar),

ncol = 1
)

```

**Comments**

+ The shading outliers for grade and condition are similar

    - it appears that both grade and condition have outliers that impact price
    - there is a steady shading pattern left to right for grade, but it is staggered
      for condition

+ Zip codes do appear to impact price

    - there are som zipcodes that have only low price homes, and some with high price
    -the slop of the regression line is weak and insignificant for zip code

**Questions**

+ Look at zip code impact on quality and condition

    - is condition and grade spread evenly between zip codes?
  
```{r}
# Show code
grid.arrange(
  # condition and grade
  houses %>% 
    ggplot(mapping = aes(x = condition, y = grade)) +
      geom_boxplot() +
    coord_flip(),
  # price and condition
  houses %>% 
    ggplot(mapping = aes(x = condition, y = price)) +
      geom_boxplot() +
    coord_flip() +
    scale_y_continuous(labels = scales :: dollar),

ncol = 1
)

```

**Comments**

+ Condition Vs. Grade

    - The distribution of grade is very focused between 5-10, with a handful of exceptions on either side

+ condition Vs. Price
    
    - The price is strung out heavily with outliers to the right, and is focused on prices around $300k-$600k
  
**Questions**

+ What impact does condition and grade have

    - are there certain zip codes with low grade, high price?
    - what is more impactful on price, grade or condition?

# Detailed EDA Questions Raised in Base EDA 7
```{r}
# Group the median expenditure and age in same graph
grid.arrange(
  
# Median price by condition and grade
houses %>%
  group_by(grade, condition) %>% 
  summarise(med_price = median(price)) %>% 
  ggplot(mapping = aes(x = grade, y = med_price, fill = condition)) + 
    geom_bar(stat = "identity", position = "dodge") +
    coord_flip() +
    scale_y_continuous(labels = scales :: dollar), # Use coord_flip() to make easier to read

# Median year built by condition and grade
houses %>%
  group_by(grade, condition) %>% 
  summarise(med_sqft= median(sqft_living)) %>% 
  ggplot(mapping = aes(x = grade, y = med_sqft, fill = condition)) + 
    geom_bar(stat = "identity", position = "dodge") +
    coord_flip(), 

ncol = 1
)

```

**Comments**

+ Median Price

    - There are Certainly homes that are very expensive that have declined in quality
    - There are also homes with poor quality that have improved, causing price to rise
    - There are large homes with both varieties of quality

**Questions**

+ Curious as to how degraded condition impacts price per foot

    - are there homes that were high quality, and are expensive with low quality condition?
    - how does location play into all of this?




## Median price and Quality 7.1

```{r}
# Group the median expenditure and age in same graph

grid.arrange(
  
# Median price by condition and grade
houses %>%
  group_by(grade, zip_code) %>% 
    summarise(med_sqft = median(sqft_living)) %>% 
    ggplot(mapping = aes(x = grade, y = med_sqft, fill = zip_code)) + 
      geom_bar(stat = "identity", position = "dodge"),# +
    #coord_flip(), # Use coord_flip() to make easier to read

# Median year built by condition and grade


houses %>%
  group_by(grade, condition) %>% 
  summarise(med_sqft= median(sqft_living)) %>% 
  ggplot(mapping = aes(x = grade, y = med_sqft, fill = condition)) + 
    geom_bar(stat = "identity", position = "dodge") +
    coord_flip(), 

houses %>%
  mutate(living_sqft_price = price/sqft_living) %>% 
  group_by(grade, condition) %>% 
  summarise(med_price_per_sqft= median(living_sqft_price)) %>% 
  ggplot(mapping = aes(x = grade, y = med_price_per_sqft, fill = condition)) + 
    geom_bar(stat = "identity", position = "dodge") +
    coord_flip(), 

ncol = 1
)

```
**Comments**

+ so it appears that the initial quality of the build does somewhat trend with the location and size.

    -there appear to be a lot more lighter blue zipcodes (98150) that are initially built very nicely
    -and it seems that most small homes had initial poor build quality, which is an interesting thought.
  
+ For homes with a median square footage of around 2,500 feet or less condition is better

    - it appears that home deterioration can happen in large homes with high quality initial builds,
    perhaps more easily than a mid size home that is easier to maintain.
    - It appears that for grade 6 homes larger than about 2,600 feet, they deteriorate a lot. 
  
+ With Median Price per squar foot in mind, it appears that low grade homes are more expensive.

    - this is could be a misinterpretation of the data, but it appears that homes that are built in low
    quality are actually more expensive per square foot.
    - this could be an inefficiency with the building process, where there is a sort of threshold to breach
     when buliding the home, that no matter the quality, it is expensive.
  


**Questions**

+ Need to examine why median price per square foot seems to be higher in poor grade homes?
  
    -is this a data error? Or is it generally more expensive per square foot in low grade initial builds?
    -medium grade homes can deteriorate to low condition quality homes.
    - Why are some high grade homes cheaper to make per square foot?
    - Does zip code play a part in price per squarefoot, like expensive area, low quality home?
    - new graph to filter out and focus on low grade homes (5 or less rating), how many observations are there?
    - compare this to high grade 10 or greater
  
## Observations of grade filtered, and review of outliers 7.2   
```{r}
#quality homes are focused in 4 zip codes, and that they all are rather expensive per sqft! Maybe some filtering of zipcodes
#here too? or a line graph?
# filter down initial grade
grade_filtered <-  c(1, 2, 3, 4, 10, 11, 12, 13)
factor(condition, levels = c(1, 2, 3, 4, 5))


grid.arrange(

houses %>%
  filter(grade %in% grade_filtered) %>%
  mutate(living_sqft_price = price/sqft_living) %>% 
  group_by(grade) %>% 
    summarise(med_price_per_sqft= median(living_sqft_price)) %>% 
    ggplot(mapping = aes(x = grade, y = med_price_per_sqft)) + 
      geom_bar(stat = "identity", position = "dodge", fill = "light blue"),

ggplot(data = houses, mapping = aes(grade)) + geom_bar(fill = "blue"), 

ggplot(data = houses, mapping = aes(condition)) + geom_bar(fill = "blue"), 

ncol = 3
)
```

**Comments**

+ The majority of the grade and condition observations are in the middle
    
    - This means that the odd interactions we keep seeing on the lower end could be extreme outliers

## Condition, Median Price per foot, By year built 7.3
```{r}
condition_by_year <- houses %>%
  mutate(living_sqft_price = price/sqft_living) %>% 
  group_by(yr_built, condition) %>% 
  summarise(med_price_per_sqft= median(living_sqft_price)) %>% 
  ggplot(mapping = aes(x = yr_built, y = med_price_per_sqft, fill = condition)) + 
    geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Year Built", y = "Median Price Per Square Foot", title = "Median Price Per Square Foot, In Respect To The Year The Home Was Built", subtitle = "Price per foot is varied, but almost feels to be trending down, with interesting outliers")
condition_by_year

ggsave(filename = "ConditionYrBuilt.png", plot = condition_by_year)

```
**comments**

+ Condition can remain excellent despite age
  
    - there are some extremely old homes that have very good condition
    - there are some very new homes that are in poor condition
    - there are homes that are old, in terrible condition, and sold for a lot
**Quesitons**
+ things to consider, why did these old homes sell for a lot in poor condition?

    - is it due to value of location?
    - is ther historical value to the homes?

## Filtered Grades with Median Price Per Square Foot of Living Space 7.4
```{r}
grade_filtered <-  c(1, 2, 3, 4, 10, 11, 12, 13)
factor(condition, levels = c(1, 2, 3, 4, 5))

grid.arrange(
#houses %>%
#  filter(grade %in% grade_filtered) %>%
#  mutate(living_sqft_price = price/sqft_living) %>% 
#  group_by(condition) %>% 
#    summarise(med_price_per_sqft= median(living_sqft_price)) %>% 
#    ggplot(data = houses, mapping = aes(x = condition, y = living_sqft_price, fill = grade)) + 
#      geom_bar(stat = "identity", position = "dodge"),   
houses %>%
  mutate(living_sqft_price = price/sqft_living) %>% 
  group_by(condition, grade) %>% 
    summarise(med_price_per_sqft= median(living_sqft_price)) %>% 
    ggplot(mapping = aes(x = condition, y = med_price_per_sqft, fill = grade)) + 
      geom_bar(stat = "identity", position = "dodge"),

Expensive_low <- houses %>%
  filter(grade %in% grade_filtered) %>%
  mutate(living_sqft_price = price/sqft_living) %>% 
  group_by(condition, grade) %>% 
    summarise(med_price_per_sqft= median(living_sqft_price)) %>% 
    ggplot(mapping = aes(x = condition, y = med_price_per_sqft, fill = grade)) + 
      geom_bar(stat = "identity", position = "dodge", guide = guide_legend(reverse = TRUE)) + theme_classic() + scale_y_continuous(labels = scales :: dollar) +
  labs(x = "Condition Ranking", y = "Median Price Per Square Foot", title = "Condition Over Time, In Respect To Cost Per Foot of Living Space", subtitle = "House condition has improved, declined and remained consistent. Price per foot is varied."), 

ncol = 2
)
ggsave(filename = "ExpensiveLowGrade.png", plot = Expensive_low) 
```
**Comments**

+ so it appears that the initial quality of the build does somewhat trend with the location and size.

  -there appear to be a lot more lighter blue zipcodes (98150) that are initially built very nicely
  -and it seems that most small homes had initial poor build quality, which is an interesting thought.
  
+ For homes with a median square footage of around 2,500 feet or less condition is better

  - it appears that home deterioration can happen in large homes with high quality initial builds,
    perhaps more easily than a mid size home that is easier to maintain.
  - It appears that for grade 6 homes larger than about 2,600 feet, they deteriorate a lot. 

+ With Median Price per squar foot in mind, it appears that low grade homes are more expensive.

  - this is could be a misinterpretation of the data, but it appears that homes that are built in low
    quality are actually more expensive per square foot.
  - this could be an inefficiency with the building process, where there is a sort of threshold to breach
   when buliding the home, that no matter the quality, it is expensive.
  



## Expense Per Living Square Foot by Zip Code 7.5
```{r}
Expensive_zips <- houses %>%
  group_by(zip_code, condition) %>% 
  mutate(living_sqft_price = price/sqft_living) %>% 
  summarise(med_price_per_sqft= median(living_sqft_price)) %>% 
  ggplot(mapping = aes(y = med_price_per_sqft, x = zip_code, fill = condition)) + 
    geom_bar(stat = "identity", position = "dodge", guide = guide_legend(reverse = TRUE)) + theme_classic() + scale_y_continuous(labels = scales :: dollar) +
  labs(x = "Zip Codes", y = "Median Price Per Square Foot", title = "There are neighborhoods that have expensive hoems", subtitle = "Our potential outliers seem to be congregated in two zip code, 98000 & 98100")

Expensive_zips

ggsave(filename = "ExpensiveZipCodes.png", plot = Expensive_zips) 


```
**Comments**

+There are certainly a large number of neighborhoods that have low condition homes

    -Of these neighbor hoods there are a few that are particularly expensive.
    -98000 and 98100 have a very odd stack of expensive spaced homes in poor condition
    
## Minor Regression Analysis 7.6
```{r}
houses <- read.csv("qp1_data.csv")#read data into environment and assign to variable

houses <- houses %>% 
  mutate(living_sqft_price = price/sqft_living)

houses <- houses %>% 
  mutate(lot_sqft_price = price/sqft_lot)

sqft_price_grade <- lm(formula = living_sqft_price ~ grade + condition + floors + bathrooms + bedrooms + zip_code + 
                         yr_built, data = houses)


summary(sqft_price_grade)

res_price <- resid(sqft_price_grade)

houses %>% 
  ggplot(mapping = aes(x = condition, y =living_sqft_price)) +
  geom_point() + 
  geom_smooth(method = lm)

Price_residuals <- ggplot(sqft_price_grade, aes(.fitted, .resid)) +
  geom_point(aes(y = res_price), color = "cadetblue4") +
  geom_hline(yintercept = 0, col = "red", linetype = "dashed") +
  labs(x = "Fitted",
       y = "Residuals",
       title = "Residuals vs Price Per Square Foot",
       subtitle = "lm(formula = living_sqft_price ~ grade + condition + floors + bathrooms + bedrooms, data = houses)") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = ),
        plot.title = element_text(hjust = 0.5, size = 20, face = "bold",
                                  color = "dodgerblue4"),
        plot.subtitle = element_text(hjust = 0.5, size = 10, color = "slategrey"),
        plot.caption = element_text(color = "gray60"))
Price_residuals

ggsave(filename = "PriceResiduals.png", plot = Price_residuals) 
```


**Comments**

+ Looking at regression analysis seems to show generalized outliers
    -I took the majority of the variables that had enough meaningful observations and used them to create a regression
    - the fitted vs. residuals is what I wanted to get at, as it indicates how the outliers look with the regression as baseline
    -From this it is easy to see that there are a dozen or so very out of range values that would cause the plots to appear
    the way they have when comparing quality vs. cost of square footage.    